---
title: "I 94 Reliability"
output:
  html_document:
    df_print: paged
---

```{r setup,message=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/dgallen/Desktop/R/Reliability')

##below are a list of packages required to run the markdown file
library(tidyverse)
library(lubridate)
library(gridExtra)
library(reshape2)
library(zoo)
library(rgdal)
library(mgcv)
library(data.table)
library(plotly)
library(broom)
library(nlstools)
library(purrr)

```




## Segment 1 N 

Read data from directory 'C:/Users/dgallen/Desktop/R/Reliability'

```{r read data, include = FALSE}
dat <- read.csv('N_1_final_rev2.csv',stringsAsFactors = FALSE)
# dat <- read.csv('N_2_final_rev3.csv',stringsAsFactors = FALSE)
# dat <- read.csv('S_3_final_rev3.csv',stringsAsFactors = FALSE)
# dat <- read.csv('S_4_final_rev2.csv',stringsAsFactors = FALSE)
# dat <- read.csv('3_N.csv',stringsAsFactors = FALSE)
# dat <- read.csv('4_N.csv',stringsAsFactors = FALSE)
# dat <- read.csv('1_S.csv',stringsAsFactors = FALSE)
# dat <- read.csv('2_S.csv',stringsAsFactors = FALSE)
```


```{r fix column names, include = FALSE}

colnames(dat) <-trimws(colnames(dat))
colnames(dat)[1] <- "TimeStamp"

```


```{r set fixed values, include=FALSE}

# 1N
Dist1 <- 2.18
Exist1 <- 129000
Lanes <- 4


#2N
# Dist1 <- 2.88
# Exist1 <- 121310
# Lanes <- 5

#3S
# Dist1 <- 2.88
# Exist1 <- 121793
# Lanes <- 3

# 4S
# Dist1 <- 2.18
# Exist1 <- 12900
# Lanes <- 4

# 3N
# Dist1 <- 1.95
# Exist1 <- 62579
# Lanes <- 3

# 4N
# Dist1 <- 1.6
# Exist1 <- 56000
# Lanes <- 2.5

# 1S
# Dist1 <- 1.6
# Exist1 <- 56000
# Lanes <- 2

# 2S
# Dist1 <- 1.95
# Exist1 <- 60645
# Lanes <- 3

head(dat)
```

### Set fixed values. 

Distance of segment is `r Dist1`
Existing Demand is `r Exist1`
Lanes in segment: `r Lanes`



### Add variables to the existing dataset. 

15 minute volume takes the VMT for each time bin and divides by the length of the segment. Unit: Vehicles per mile.
Speed is the segment length divided by the travel time ( * 60). Unit: Miles per hour.
Hourly_Flow is the 15 minute volume multipled by 4 and divided by the number of lanes. Unit: Vehicles per hour.
Density is the Hourly Flow divided by speed. Unit: Vehicles per hour.

We also create a variable that measures the delay type experienced by the segment. Weather (snow) events and crashes were tracked for the time period. If the segment experiences a weather event and a crash, weather will be the controlling delay type.

```{r add_variables}

Rel_2014 <- dat %>% mutate(Vol_15_min  = VMT_total/Dist1,
                           Speed       = Dist1/TT_mean*60,
                           Hourly_Flow = Vol_15_min*4/Lanes,
                           Dens        = Hourly_Flow/Speed,
                           DateTime    = as.POSIXct(TimeStamp,format="%m/%d/%Y %H:%M"),
                           # DateTime    = as.POSIXct(TimeStamp,format="%Y%m%d %H:%M"),
                           YearDay     = weekdays(DateTime),
                           DateTimeSec = as.numeric(DateTime),
                           Delay_type  = (ifelse(weather == 1, 
                                                'Weather',
                                                ifelse(Crash == 1, 
                                                       'Crash',
                                                       'None'))),
                           Index = seq(1,length(Speed),1)) %>%
  filter(!is.na(DateTime))
```

```{r filter for outlier}
boxplot(Rel_2014$Dens)

Rel_2014 <- Rel_2014 %>%
  mutate(Rollavg = (rollmean(TT_mean,k=8,fill = TT_mean, align = 'right') +
                      rollmean(TT_mean,k=8,fill = TT_mean, align = 'left'))/2) %>%
  mutate(outlier = ifelse(abs(Rollavg - TT_mean) > 6 * sd(Rollavg), TRUE, FALSE))  
  # filter(Dens <= 100) #3N


```

``` {r bin data}

# 1N
# bins <- c(seq(floor(min(Rel_2014$Dens)),30,2),
#           40,50,ceiling(max(Rel_2014$Dens)))


bins <- c(seq(floor(min(Rel_2014$Dens)),max(Rel_2014$Dens)+2,2))

# bins <- c(seq(floor(min(Rel_2014$Dens)),24,2),
#           max(Rel_2014$Dens)+1)

Rel_2014$cut <- cut(Rel_2014$Dens,bins,include.lowest = TRUE)
Rel_2014$Density_bin <- as.numeric(stringi::stri_match_last_regex(Rel_2014$cut,'[0-9]+')) 

Rel_2014 %>%
  group_by(Density_bin) %>%
  summarise(Count = n()) %>%
  arrange(Density_bin)
```




### Percentile Bins

We will break down the dataset into each delay type and then further into percentiles. 

```{r percentiles, include=FALSE}
p <- c(.05,.1,.25,.5,.75,.9,.95)
```


The percentiles choosen are `r p`.

For each quantile a travel time quantile is calculated. We also want to calculate a Free Floew Speed for each quantile. 

```{r create binned dataset, include=FALSE}

Rel_2014_bin_dt <- Rel_2014 %>%
  mutate(Density_bin = as.numeric(stringi::stri_match_last_regex(cut,'[0-9]+'))) %>%
  group_by(Delay_type,Density_bin) %>%
  summarise(quantiles = list(sprintf("%1.0f%%", p*100)),
            TT_quantile = list(quantile(TT_mean,p))) %>%
  unnest() %>%
  mutate(Speed = Dist1/TT_quantile*60,
         Flow = Speed * Density_bin,
         quantiles = as.factor(quantiles)) %>%
  as.data.frame()

Rel_2014_FF <- Rel_2014_bin_dt %>%
  group_by(Delay_type,quantiles) %>%
  summarise(FF_quantile = round(quantile(Flow/Density_bin,.95, na.rm = TRUE),0))


Rel_2014_bin_dt <- merge(Rel_2014_bin_dt,Rel_2014_FF) %>%
  mutate(Index_bin = as.integer((seq(1,length(FF_quantile),1))),
         TTI = TT_quantile/(Dist1/FF_quantile * 60))

head(Rel_2014_bin_dt)
```

The Free Flow speeds were calcuated using the 95th percentile speeds. 

```{r, fig.width = 11}
ff <- ggplot()+
  geom_point(data = data.frame(Rel_2014_bin_dt),aes(x=Density_bin,y=Flow, color=quantiles))+
  geom_abline(data = data.frame(Rel_2014_FF), aes(slope = FF_quantile, intercept = 0, linetype = 'dashed', color = quantiles))+
  facet_grid(~Delay_type)
ggplotly(ff)
```


A look at the Density vs travel time plots. A generic smooting function has been added. 

```{r Density vs TT, fig.width = 11}
tt <- ggplot()+
  geom_point(data=Rel_2014_bin_dt,aes(x= Density_bin,y = TT_quantile, color= quantiles))+
  geom_smooth(data=Rel_2014_bin_dt,aes(x= Density_bin,y = TT_quantile, color= quantiles),se=F)+
  facet_grid(~Delay_type)
ggplotly(tt)
```


We need to find the congestion point for each Delay type and percentile.

```{r find congestion points, include=FALSE}

model_crash <- Rel_2014_bin_dt %>%
  filter(Delay_type == 'Crash') %>%
  nest(-Delay_type,-quantiles) %>%
  # as.tibble() %>%
  mutate(fit = map(data, ~ loess(Flow ~ Density_bin, span = 0.75, data = .,)),
         results = map(fit, augment)) %>% 
  as.tibble() %>%
  unnest(results)

model <- Rel_2014_bin_dt %>%
  filter(Delay_type != 'Crash') %>%
  nest(-Delay_type,-quantiles) %>%
  # as.tibble() %>%
  mutate(fit = map(data, ~ loess(Flow ~ Density_bin, span = 1, data = .,)),
         results = map(fit, augment)) %>% 
  as.tibble() %>%
  unnest(results) %>%
  rbind(model_crash)

model_max <- model %>%
  group_by(Delay_type,quantiles) %>%
  # summarise(max = max(.fitted))
  filter(.fitted == max(.fitted)) %>%
  select(Delay_type, quantiles, Cong_Flow = .fitted, Cong_Density = Density_bin)
```

### Plots

Below are the Flow vs Density plots for each condition. 

Verical lines show the congestion Density for each condition. 

```{r plots weather delay,  fig.width = 10, fig.height=5}


f <- ggplot()+
  geom_point(data=filter(Rel_2014_bin_dt,Delay_type == 'Weather') ,
             aes(x= Density_bin,y = Flow, color= quantiles))+
  geom_smooth(data=filter(Rel_2014_bin_dt,Delay_type == 'Weather'),
              aes(x= Density_bin,y = Flow, color= quantiles),span =1,se=F)+
  geom_vline(data=filter(model_max,Delay_type == 'Weather'),aes(xintercept=Cong_Density, color = quantiles)) +
  ggtitle('Delay Type = "Weather"')
ggplotly(f)

```


```{r plots crash delay,  fig.width = 10, fig.height=5}


f <- ggplot()+
  geom_point(data=filter(Rel_2014_bin_dt,Delay_type == 'Crash') ,
             aes(x= Density_bin,y = Flow, color= quantiles))+
  geom_smooth(data=filter(Rel_2014_bin_dt,Delay_type == 'Crash'),
              aes(x= Density_bin,y = Flow, color= quantiles),span =0.75,se=F)+
  geom_vline(data=filter(model_max,Delay_type == 'Crash'),aes(xintercept=Cong_Density, color = quantiles)) +
  ggtitle('Delay Type = "Crash"')
ggplotly(f)

```

```{r plots no delay, fig.width = 10, fig.height=5}


f <- ggplot()+
  geom_point(data=filter(Rel_2014_bin_dt,Delay_type == 'None') ,
             aes(x= Density_bin,y = Flow, color= quantiles))+
  geom_smooth(data=filter(Rel_2014_bin_dt,Delay_type == 'None'),
              aes(x= Density_bin,y = Flow, color= quantiles),span =1,se=F)+
  geom_vline(data=filter(model_max,Delay_type == 'None'),aes(xintercept=Cong_Density, color = quantiles)) +
  ggtitle('Delay Type = "None"')
ggplotly(f)
```


### Demand

We now calcualte Demand.

If Density is less than the congestion density, Demand = Flow.
If density is greated that the congestion density, Demand = 2 x Congestion Flow - Flow. 

TTI is the travel time divided by the travel time at Free Flow. 




```{r calculate demand}
Rel_2014_bin_dt <- merge(Rel_2014_bin_dt,model_max,by = c('Delay_type','quantiles')) %>%
  mutate(Demand = ifelse(Density_bin <= Cong_Density,Flow,2*Cong_Flow - Flow),
         DC = Demand/Cong_Flow,
         TTI_mod = TTI - 1)

```

```{r demand vs TTI,  fig.width = 10, fig.height=5}

s <- ggplot()+
  # geom_point(data=Rel_2014_bin_dt,aes(x= Demand,y = TTI_pred, color= quantiles))+
  geom_smooth(data=Rel_2014_bin_dt,aes(x= Demand,y = TTI, color= quantiles),se=F)+
  geom_point(data=Rel_2014_bin_dt,aes(x= Demand,y = TTI, color= quantiles))+
  # geom_smooth(se=F)+
  # geom_abline(data=Rel_2014_FF,aes(slope = FF,intercept = FF,colour = as.factor(quantiles)))+
  facet_grid(~Delay_type)
ggplotly(s)

```


## Model

We now fit a curve to model the demand. 

On the x axis we plot Demand/Congestion Flow and on the Y axis we plot TTI.

The formula we use to model the data is 1 + aplha x D/C ^ beta. A set initial portion of each curve is set to TTI = 1, to correctly model Demand = Flow for now congestion conditions. 

Below is the model output for no delay conditions, 50th percentile:

```{r develop model}
sample_data <- Rel_2014_bin_dt %>%
  filter(Delay_type == 'Crash',
         quantiles == '95%') %>%
  select(DC, TTI)


formulaExp <- as.formula(TTI ~(DC <= .95) + (DC > .95) * ( mu * (DC ^ b)))
b <- 3
mu <- 1.5
# preview(formulaExp, data = sample_data, start = list(mu, b))

sample_nls <- nls(formulaExp, data = sample_data, start = list(mu = 1.5, b = 3))
plotfit(sample_nls, smooth = TRUE)

pred <- data.frame(DC = 0.99)
predict(sample_nls,pred)
```


```{r, include=FALSE}

# 2N
# breaks <- model_max %>% data.frame() %>%
#   mutate(br = c(0.86,0.86,0.87,0.865,0.865,0.865,0.865,
#                 0.88,0.895,0.895,0.895,0.95,1.04,1.03,
#                 0.91,0.9,0.91,0.91,.89,0.86,0.79)) %>%
#   select(Delay_type,quantiles,br)

# 3S
# breaks <- model_max %>% data.frame() %>%
#   mutate(br = c(0.92,0.91,0.91,0.92,0.87,0.85,0.92,
#                 0.87,0.9,0.86,0.9,0.9,0.91,0.9,
#                 0.5,0.48,0.5,0.48,0.53,0.49,0.48)) %>%
#   select(Delay_type,quantiles,br)

# 1N
breaks <- model_max %>% data.frame() %>%
  mutate(br = c(0.87,0.88,0.87,0.89,0.91,0.91,.845,
                0.94,.94,0.95,0.96,0.99,0.99,0.99,
                0.99,0.99,0.99,0.98,0.96,0.96,0.95)) %>%
  select(Delay_type,quantiles,br)

Rel_2014_bin_dt <- merge(Rel_2014_bin_dt,breaks)
```

```{r Demand model test, include=FALSE}

func_exp <- function(x) {
  nls(
    TTI ~(DC <= br) + (DC > br) * ( mu * (DC ^ b)),
    start = list(mu = 1.5, b = 3),
    data = x)
}


formulaExp <- as.formula(TTI ~(DC <= br) + (DC > br) * ( mu * (DC ^ b)))

model_nls <- Rel_2014_bin_dt %>%
  nest(-Delay_type,-quantiles) %>%
  mutate(fit = purrr::map(data, func_exp)) 
  # mutate(pred = purrr::map2(fit,data,predict)) %>%
  # unnest(data,pred)

```

### Predict

We join each model back to the original dataset and create a predicted TTI for each row. 

```{r join roll up, include=FALSE}

Rel_2014_TT_dt <- Rel_2014 %>%
  select(Delay_type,Density_bin,TT_quantile = TT_mean, Index) %>% 
  data.table()


Rel_2014_bin_dt <- data.table(Rel_2014_bin_dt)
setkey(Rel_2014_bin_dt, NULL)
setkey(Rel_2014_bin_dt,Delay_type,Density_bin,TT_quantile)
join_TT <- Rel_2014_bin_dt[Rel_2014_TT_dt, roll = 'nearest', .(TT_q=x.TT_quantile,
                                                 TT_mean = i.TT_quantile,
                                                 Density_bin = Density_bin,
                                                 Delay_type = Delay_type,
                                                 Quantiles = quantiles,
                                                 FF = FF_quantile,
                                                 Cong_Flow = Cong_Flow,
                                                 Cong_Density = Cong_Density,
                                                 Index = Index,
                                                 br = br)]

join_TT <- join_TT[!duplicated(join_TT$Index),]


Rel_2014_f <- merge(Rel_2014, join_TT[,c('Index','TT_q','Quantiles','Cong_Flow','Cong_Density','FF','br')], by = 'Index', all.x=TRUE)
table(is.na(Rel_2014_f$Cong_Flow))

Rel_2014_f <- Rel_2014_f %>%
  mutate(Demand  = ifelse(Density_bin <= Cong_Density,Hourly_Flow,2*Cong_Flow - Hourly_Flow),
         DC = Demand/Cong_Flow,
         TTI = TT_mean/(Dist1/FF * 60)) %>%
  nest(-Delay_type, -Quantiles)

Rel_2014_f <- merge(Rel_2014_f,model_nls %>% select(Delay_type, Quantiles = quantiles, fit)) %>%
  as.tibble() %>%
  mutate(pred = map2(fit,data,predict)) %>%
  as.tibble() %>%
  unnest(data,pred)
```

```{r plot pred vs TTI, fig.width = 10, fig.height=5}

pr <- ggplot(Rel_2014_f, aes(x=TTI,y=pred, color=Quantiles))+
  geom_point()+
  facet_grid(~Delay_type)
ggplotly(pr)

```



```{r alternative models, include = FALSE}


# model_demand <- Rel_2014_bin_dt %>%
#   nest(-Delay_type,-quantiles) %>%
#   as.tibble() %>%
#   mutate(fit = map(data, ~ lm(log(TTI_mod) ~ log(DC), data = .)),
#          results = map(fit, tidy)) %>% 
#   as.tibble() %>%
#   unnest(results) %>%
#   dcast(Delay_type + quantiles ~ term, value.var = 'estimate') %>%
#   mutate(Alpha = exp(`(Intercept)`))
#   #        TT = exp(log.TTI_mod.)+1,
#   #        Demand = exp(log.DC.)*Cong_Density)
# 
# 
# 
# model_demand_aug <- Rel_2014_bin_dt %>%
#   nest(-Delay_type,-quantiles) %>%
#   as.tibble() %>%
#   mutate(fit = map(data, ~ lm(log(TTI_mod) ~ log(DC), data = .)),
#          results = map(fit, augment)) %>% 
#   as.tibble() %>%
#   unnest(results) 


# Rel_2014_bin_dt <- merge(Rel_2014_bin_dt,model_demand[,c('log.TTI_mod.','log.DC.','Index_bin')])
# Rel_2014_bin_dt <- merge(Rel_2014_bin_dt,model_demand)
# 
# Rel_2014_bin_dt$TTI_pred <- 
#   Rel_2014_bin_dt$Alpha * (Rel_2014_bin_dt$DC ^ Rel_2014_bin_dt$`log(DC)`) + 1

# 
# Rel_2014_FF_dt <- Rel_2014 %>%
#   select(Delay_type,Density_bin,Hourly_Flow, Index) %>% 
#   data.table()
# 
# setkey(Rel_2014_bin_dt,Delay_type,Density_bin,Hourly_flow)
# join_FF <- Rel_2014_bin_dt[Rel_2014_FF_dt, roll = -Inf, .(Flow_q=x.Flow_quantile,
#                                                  Hourly_flow = i.Hourly_Flow,
#                                                  Density_bin = Density_bin,
#                                                  Delay_type = Delay_type,
#                                                  Quantiles = quantiles,
#                                                  FF = FF_quantile,
#                                                  Index = Index)]
# 
# 
# Rel_2014 <- merge(Rel_2014, join[,c('Index','FF')], by = 'Index', all.x=TRUE)
# table(is.na(Rel_2014$FF))
```

```{r bin demand, include=FALSE}

# Rel_2014 <- merge(Rel_2014, Rel_2014_FF, by = c('Delay_type'))



# bins_demand <- c(floor(min(Rel_2014$Demand)),
#           seq(400,1900,100),
#           seq(2000,3800,200),
#           seq(4000,6000,1000),9000,ceiling(max(Rel_2014$Demand)))
# 
# Rel_2014$cut_demand <- cut(Rel_2014$Demand,bins_demand,include.lowest = TRUE, dig.lab = 10)
# Rel_2014$Demand_bin <- as.numeric(stringi::stri_match_last_regex(Rel_2014$cut_demand,'[0-9]+')) 
# 
# Rel_demand_bin_dt <- Rel_2014 %>%
#   group_by(Delay_type, Demand_bin) %>%
#   summarise(quantiles = list(sprintf("%1.0f%%", p*100)),
#             TT_quantile = list(quantile(TT_mean,p)),
#             Count = n()) %>%
#   unnest() %>%
#   mutate(quantiles = as.factor(quantiles)) %>%
#   select(Demand_bin,Delay_type,TT_mean = TT_quantile,quantiles,Count) %>%
#   data.table()
# 
# 
# Rel_2014_dt <- Rel_2014 %>%
#   select(Demand_bin, Delay_type,TT_mean, Index) %>% 
#   data.table()
```

```{r join demand, include=FALSE}
# setkey(Rel_demand_bin_dt,Demand_bin,Delay_type,TT_mean)
# join_demand <- Rel_demand_bin_dt[Rel_2014_dt, roll = 'nearest', .(TT=x.TT_mean,
#                                                  TT_mean = i.TT_mean,
#                                                  Demand_bin = Demand_bin,
#                                                  Delay_type = Delay_type,
#                                                  Demand_quantiles = quantiles,
#                                                  Index = Index)]
# 
# 
# Rel_2014 <- merge(Rel_2014, join_demand[,c('Index','Demand_quantiles')], by = 'Index', all.x=TRUE)
# dim(Rel_2014)
```









